---
title: "map-leaflet2"
author: "Jonas"
date: "2025-05-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
# This is a Shiny time series map web application
library(shiny)
library(tidyverse)
library(tidyr)
library(leaflet)
library(xts)
```

```{r file}
# Read the CSV file - replace with your actual file path
algae_data <- read_delim(file = "C:\\Users\\jonas\\Documents\\uni\\TM\\SP\\map-script\\obs_algues_export_2025-01.csv",
                         delim = ';',
                         locale = locale(encoding = "ASCII"))
```

```{r basic-data}
# Clean and prepare the data
# Select only the necessary columns
df <- algae_data %>%
  mutate(
    Latitude = ifelse(!is.na(`loc-point-Latitude`), 
                     `loc-point-Latitude`, 
                     `loc-point_auto_5-Latitude`),
    Longitude = ifelse(!is.na(`loc-point-Longitude`), 
                      `loc-point-Longitude`, 
                      `loc-point_auto_5-Longitude`)
  ) %>%
  # Filter out rows with missing coordinates
  filter(!is.na(Latitude) & !is.na(Longitude)) %>%
  # Select relevant columns for mapping
  select(`groupe-date_heure`, Latitude, Longitude, `obs-bloom`, `contexte-nom_lac`)

names(df) = c("Date", "lat", "lon", "bloom", "nomlac")
df$year <- lubridate::year(df$Date)
df$month <- lubridate::month(df$Date, label = TRUE, abbr = FALSE)
df$week <- lubridate::week(df$Date)
df$Date <- as.Date(df$Date)
head(df)
```

```{r example}
xts_to_tibble <- function(xts_obj) {
  data.frame(index(xts_obj), coredata(xts_obj)) %>%
    set_names(c("date", names(xts_obj))) %>%
    as_tibble()
}

ui <- fluidPage(

  # Title
  titlePanel("Time Series Visualization Map"),
  sidebarLayout(

    # Define the sidebar
    sidebarPanel(
      radioButtons(
        inputId = "Frequency",
        label = " Select Timer Series Frequency",
        choices = c(
          "weeks",
          "months",
          "years"
        ),
        selected = "weeks",
        inline = T
      ),
      uiOutput("Time_Series_UI")
    ),
    mainPanel(
      leafletOutput("Time_Series_Map")
    ),
  )
)



# Define server logic required to draw a histogram
server <- function(input, output) {

  # Render slider input depending on data frequency

  observe({
    # Create an xts object
    df_xts <- xts(df, order.by = as.Date(df$Date))

    # All_Dates = unique(df$Start_Date)

    Filtered_Dates <- df_xts[xts::endpoints(
      df_xts,
      on = input$Frequency
    )] %>% xts_to_tibble()

    output$Time_Series_UI <- renderUI({
      sliderInput("Date", "Date:",
        min = pull(slice_min(Filtered_Dates, date), date),
        max = pull(slice_max(Filtered_Dates, date), date),
        value = pull(slice_min(Filtered_Dates, date), date),
        step = 1,
        timeFormat = "%YYYY-%MM-%DD",
        animate = T
      )
    })
  })

  # Filter data for the date selected
  Filtered_Data <- reactive({
    req(input$Date)
    filter(df, Date == input$Date)
  })


  # Create the leaflet map
  output$Time_Series_Map <- renderLeaflet({
    leaflet(df) %>%
      addProviderTiles(providers$OpenStreetMap)
  })

  # Create data markers for selected date
  observe({
    leafletProxy("Time_Series_Map", data = Filtered_Data())     %>%
      clearGroup("points") %>%
      addCircleMarkers(group='points',
        lng = ~lon, lat = ~lat,
        popup = ~paste(
                    "<strong>Lac:</strong> ", nomlac
        ))
  })
}

# Run the application
shinyApp(ui = ui, server = server)  
```